Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2018-12-05 Manuscript/manuscript.Rmd
manuscript edits

The performances were found to be consistently poor for models with no re-sampling across the other resolutions (Supplementary Figure 13; Supplementary Table 10). The low, and sometimes missing, values indicate that the model is unable to identify many, if any, true positive outcomes. That is, given imbalanced data, the model was unable to predict which genomic bins truly contain a flanked TAD boundary.

@@

When evaluatating the 5kb resolution data for a randomly under-sampled training set, it was found that models obtained better classification performance for F1 and MCC metrics (Figure 14). The increase in performance was present across both overlap predictor types. Models also featured a decrease in both accuracy and AUC, which was attributed to less inflated false negative rates (Supplementary Table 11). Variable reduction was found not to have an affect on resampling, and subsequent performance as any differences were again marginal. Performances were also improved among the models using the other data resolutions as well. However, model improvement dissipated as resolution increased (Supplementary Figure 15; Supplementary Table 12).

@@

We then used the ensemble to compare models with distance type predictors for the feature space. For the 5kb resolution data, it was found that models with distance type predictors out performed models with either of the two overlap types (Figure 16). The improvement was observed across all 4 performance metrics that were considered (Supplementary Table 13). Likewise, improved performances were seen in the other 3 resolutions as well. Although overall performance decreased as resolution increased, models with distance type predictors still were able to maintain a profound improvement (Supplementary Figure 14;Supplementary Table 14).

@@

The differences in performances was most evident when analyzing the variable importance plots for each model using the 4 resolutions. We found that when we analyzed the variable importance produced by each random forest algorithm. We found that for the 5kb resolution, the most predictive elements include insulator proteins, DNase hypersensitive sites, the transcription factor CTCF, as well as histone modifications H3k9ac, H2az, and H3k4me2 (Figure 10 A). Although, the insulator protein is featured as a top predictor in 2 of the other 3 resolutions, the rest of the predictors are less consistent among each other and less in agreement with the literature (Figure 10 B-D). The results from Figures 9 and 10 confirm that the model results appear to degrade as resolution increases with respect to both prediction and the identification of genomic elements associated with TAD boundaries.

>>>>

The performances were found to be consistently poor for models with no re-sampling across the other resolutions (Supplementary Figure 13; Supplementary Table 10). The low, and sometimes missing, values indicated that the model was unable to identify many, if any, true positive outcomes. That is, given imbalanced data, the model was unable to predict which genomic bins truly overlapped with a flanked TAD boundary.

@@

When evaluatating the 5kb resolution data for a randomly under-sampled training set, it was found that models obtained better classification performance for F1 and MCC metrics (Figure 14). The increase in performance was present across both overlap predictor types. Models also featured a decrease in both accuracy and AUC, which was attributed to less inflated false negative rates (Supplementary Table 11). Variable reduction was found not to have an affect on resampling, and subsequent performances, as any differences were again marginal. Performances were also improved among the models using the other data resolutions as well. However, we found that model improvement dissipated as resolution increased (Supplementary Figure 15; Supplementary Table 12).

@@

We then used the ensemble to compare models with distance type predictors for the feature space. For the 5kb resolution data, it was found that models with distance type predictors out performed models with either of the two overlap types, regardless of the regularization technique performed (Figure 16). The improvement was observed across all 4 performance metrics that were considered (Supplementary Table 13). Likewise, improved performances were seen in the other 3 resolutions as well. Although overall performance decreased as resolution increased, models with distance type predictors still were able to maintain a profound improvement (Supplementary Figure 14; Supplementary Table 14).

@@

The differences in performances was most evident when comparing the variable importances across the 4 resolutions. We found that for the 5kb resolution, the most predictive elements included insulator proteins, DNase hypersensitive sites, the transcription factor CTCF, as well as histone modifications H3k9ac, H2az, and H3k4me2 (Figure 10 A). Although, the insulator protein is featured as a top predictor in 2 of the other 3 resolutions, the rest of the predictors were less consistent among each other and less in agreement with the literature (Figure 10 B-D). The results from Figures 9 and 10 confirm that the model results appear to degrade as resolution increases with respect to both prediction and the identification of genomic elements associated with TAD boundaries.
